---
title: "Get Data"
author: "&copy;  2023 Vishal Lala"
format:
  html:
    toc: true
    toc-depth: 3
    toc-float: true
    theme: cosmo
    highlight-style: ally
    df-print: paged
    embed-resources: true
  docx:
    toc: true
    toc-depth: 3
    highlight-syle: ally
  pdf:
    toc: true
    toc-depth: 3
    documentclass: scrartcl
    papersize: letter
    highlight-syle: ally
execute:
  cache: true
editor: 
  markdown: 
    wrap: 72
---

# Getting Data

1.  Download

2.  read

3.  API

4.  Scrape

## Download

Simplest way to get data is to get data manually (e.g., web download,
email attachment, disk) and then view it.

### NYPD Data

NYPD Personnel Demographics data have been compiled on [Open
NYC](https://data.cityofnewyork.us/Public-Safety/NYPD-Personnel-Demographics/5vr7-5fki/data).
Download manually and then read or read directly from website.

```{r}
nypd = read.csv('https://data.cityofnewyork.us/api/views/5vr7-5fki/rows.csv?accessType=DOWNLOAD')
```

#### Explore the data

What are the names and types of variables included in this dataset?

```{r}
table(nypd$YearsOnJob..bins.)
```

How many rows and columns are in this dataset?

```{r}

```

Are the variables sex and gender the same?
```{r}

```


Are women well-represented on the NYPD? Specifically, what percent of NYPD officers are female?

```{r}

```

What is the proportion of officers by race?
```{r}

```
Let us examine the trend in gender hiring based on experience (i.e., years on the job). To make the results easier to examine, we will put the data in a wide format using library(tidyr) function pivot_wider(). 
```{r}
library(tidyr)
nypd |>
  group_by(YearsOnJob..bins., Gender)|>
  summarize(n = n())|>
  pivot_wider(names_from = Gender,values_from = n)
```


### Beach Water Quality

Open NYC regularly updates quality of water in New York city beaches and
posts them to [NYC Open Data](https://data.cityofnewyork.us/Health/DOHMH-Beach-Water-Quality-Data/2xir-kwzz).
Data can be downloaded manually and then read. Or the steps can be automated by directly reading from the site.

```{r}
water_quality = read.csv("C:/Users/vlala/Downloads/DOHMH_Beach_Water_Quality_Data.csv")
str(water_quality)
```

What is the highest level of Enterocci recorded?

```{r}

```

Which are the Ten Worst Beaches based on levels of Enterococci?

```{r}

```
Which are the Ten Worst Beaches based on levels of Enterococci?
```{r}

```


### Baby Names

Social Security Administration shares data on names of all children born in the US. Download zip file containing State Specific names from (SSA)[https://www.ssa.gov/oact/babynames/limits.html] and extract names for New York, ny.txt.

```{r}
babies = read.csv('https://raw.githubusercontent.com/pseudorational/LubinAnalyticsBootcamp/main/NY.TXT', 
         header=F, 
         col.names = c('state','gender','year','name','n'))
head(babies)
```

Next, identify top 25 boy names since 2010

```{r}
library(dplyr); library(ggplot2); library(forcats)
boys = 
  babies%>%
  filter(year>2010)%>%
  group_by(gender, name)%>%
  summarize(total_n = sum(n))%>%
  mutate(rank = as.integer(rank(x = desc(total_n),ties.method = 'first')))%>%
  ungroup()%>%
  filter(rank<=25, gender=='M')%>%
  select(rank, name, total_n)%>%
  arrange(rank)%>%
  ggplot(aes(x=forcats::fct_reorder(name,total_n),y=total_n, fill=desc(total_n)))+
  geom_col()+
  coord_flip()+
  guides(fill=F)+
  ylab('')+xlab('Number Since 2010')+
  ggtitle('Top 25 Boy Names in New York since 2010')
boys
```

Next, identify top 25 girl names since 2010

```{r}
girls = 
  babies%>%
  filter(year>2010)%>%
  group_by(gender, name)%>%
  summarize(total_n = sum(n))%>%
  mutate(rank = as.integer(rank(x = desc(total_n),ties.method = 'first')))%>%
  ungroup()%>%
  filter(rank<=25, gender=='F')%>%
  select(rank, name, total_n)%>%
  arrange(rank)%>%
  ggplot(aes(x=forcats::fct_reorder(name,total_n),y=total_n, fill=desc(total_n)))+
  geom_col()+
  coord_flip()+
  guides(fill=F)+
  ylab('')+xlab('Number Since 2010')+
  ggtitle('Top 25 Girl Names in New York since 2010')
girls
```

### Dog Names

New York City has published a list of dog names which can be downloaded
as a csv file from (Open NYC)[https://data.cityofnewyork.us/Health/NYC-Dog-Licensing-Dataset/nu7n-tubp/data]

```{r}
dogs = read.csv('https://data.cityofnewyork.us/api/views/nu7n-tubp/rows.csv?accessType=DOWNLOAD')
str(dogs)
```

List the top 15 Male Dog names since 2010

```{r}
dogs_boy = 
  dogs%>%
  rename('name'='AnimalName','gender'='AnimalGender','year'='AnimalBirthYear')%>%
  mutate(name = stringr::str_to_title(name))%>%
  filter(!(name %in% c('Unknown','Name Not Provided','None','Name','Not','Dog')))%>%
  group_by(year, gender, name)%>%
  summarize(n = n())%>%
  filter(year>2010)%>%
  group_by(gender, name)%>%
  summarize(total_n = sum(n))%>%
  mutate(rank = as.integer(rank(x = desc(total_n),ties.method = 'first')))%>%
  ungroup()%>%
  filter(rank<=25, gender=='M')%>%
  select(rank, name, total_n)%>%
  arrange(rank)%>%
  ggplot(aes(x=forcats::fct_reorder(name,total_n),y=total_n, fill=desc(total_n)))+
  geom_col()+
  coord_flip()+
  guides(fill=F)+
  ylab('')+xlab('Number Since 2010')+
  ggtitle('Top 25 Boy DOG Names in New York since 2010')
dogs_boy
```

List the top 15 Female Dog names since 2010

```{r}
dogs_girl = 
  dogs%>%
  rename('name'='AnimalName','gender'='AnimalGender','year'='AnimalBirthYear')%>%
  mutate(name = stringr::str_to_title(name))%>%
  filter(!(name %in% c('Unknown','Name Not Provided','None','Name','Not','Dog')))%>%
  group_by(year, gender, name)%>%
  summarize(n = n())%>%
  filter(year>2010)%>%
  group_by(gender, name)%>%
  summarize(total_n = sum(n))%>%
  mutate(rank = as.integer(rank(x = desc(total_n),ties.method = 'first')))%>%
  ungroup()%>%
  filter(rank<=25, gender=='F')%>%
  select(rank, name, total_n)%>%
  arrange(rank)%>%
  ggplot(aes(x=forcats::fct_reorder(name,total_n),y=total_n, fill=desc(total_n)))+
  geom_col()+
  coord_flip()+
  guides(fill=F)+
  ylab('')+xlab('Number Since 2010')+
  ggtitle('Top 25 Girl DOG Names in New York since 2010')
dogs_girl
```

#### Do dog parents name their dogs like their children?

```{r}
library(gridExtra)
grid.arrange(boys, dogs_boy, girls, dogs_girl)
```
